%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ACKNOWLEDGEMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{ACKNOWLEDGEMENTS}
\label{sec:acknowledgements}

\addcontentsline{toc}{chapter}{ACKNOWLEDGEMENTS}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.


\indent I would like to express my sincere gratitude to my advisor Dr. Alex Sprintson for the guidance, inspiration and the mentoring he provided all throughout my graduate career. He always provided great source of encouragement and was always available to discuss various problems that eventually led to this thesis. I would also like to thank the entire Flowgrammable Team who supported and provided valuable inputs throughout the work. I greatly benefited from the discussions and hope that we continue to work together in other projects. I am also thankful to Muxi Yan for the interactions and brain storming sessions during the \aetherflow project. A special mention also goes out to Jalil Modares and Dr. Nicholas Mastronarde from Univerity at Buffalo for the collaborative effort which went into \crossflow project. Finally, I want to thank Dr. Paul Gratz and Dr. Radu Stoleru for having served as my committe members.

\indent Last but not least, I want to thanks my family and friends for their immense support and encouragement, and most importantly, I would like to thank my wife Punam for her understanding and support, without whom none of this would have been possible.


\pagebreak{}
